Testing and Isolation Efficacy: Insights from a Simple Epidemic Model

### 2021-04-08 SIR_manuscript.tex
old analysis of diff R0 wrt rho was wrong and ditched

The hard thing to prove is to show the reduction of $\Rnum$ with respect to testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, i.e., $0<\omega$, $0 \leq \rho<\omega$, and $0 \leq \theta\_w \leq \theta\_c \leq 1$.

% %%%%%%%
\subsection{More testing rate reduces $\Rnum$}

This is to prove that $\pro \Delta > 0$ with $\Delta$ defined in Eq. \eqref{eq:abc}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eqs. \eqref{R0}.
We will show that (a) $\pro (A/C)>0$ and (b) $\pro (B/C)>0$.
Note that we will use the following information: 
\begin{enumerate}
\item
$\pro \hat F_I >0$ for  all $\rho$ and specifically the feasible range, $\rho \in [0,\omega]$. Here $\hat F_I = (\omega\rho/(\omega-\rho))w_I/w_S$ for the \percap testing rate of infectious individuals near the disease-free equilibrium,
\item $\pro \hat F_I x\_u^* >0$ since $\hat F_I x\_u^* = \rho w_I/w_S$,
\item $\pro \hat F_I x\_n^* >0$ since $\hat F_I x\_n^* = \frac{\rho^2}{\omega-\rho} w_I/w_S$, 
\item $\pro x\_n^* >0$ and $\pro x\_u^* <0$ given the DFE expression in \ref{dfe},
\item
\begin{align}
\pro A &= 
\gamma \Big(\gamma(\omega+\gamma)\pro x\_n^* + \omega p_I \pro(\hat F_I x\_n^*)+(\omega+\gamma) \pro(\hat F_I)
\Big),\\
\pro B &= \gamma \omega p_I \pro(\hat F_I x\_u^*)+\omega^2 p_I \pro(\hat F_I),\\
\pro C &=(\omega+\gamma)(\omega p_I+\gamma) \pro(\hat F_I).
\end{align}
\end{enumerate}

(a) $\pro (A/C)>0$;

\begin{align}
& \pro (A/C) = \pro{A} \dot C - \pro{C} \dot A \\
&= \gamma (\omega+\gamma)^2(\omega p_I+\gamma) 
\Big( 
[\gamma \pro x\_n^* + \frac{\omega p_I}{(\omega+\gamma)} \pro(\hat F_I x\_n^*) + \pro\hat F_I]
[\gamma\frac{(\omega+\gamma)}{(\omega p_I+\gamma)}+\hat F_I] - \\
& \pro(\hat F_I)
[\gamma x\_n^* + \frac{\omega p_I}{(\omega+\gamma)}(\hat F_I x\_n^*) + \hat F_I] 
\Big) 
\end{align}

Since $\pro(x\_n^*) >0$ and $\pro (\hat F_I x\_n^*)>0$,  it is enough to show that the coefficient of 
$\pro (\hat F_I)$ is summed up to a positive value.
$\pro (\hat F_I)$ coefficient is as follows
$$\big( \frac{\omega p_I}{(\omega+\gamma)} x\_n^* +1 \big) 
\big( \gamma\frac{(\omega+\gamma)}{(\omega p_I+\gamma)} + \hat F_I \big) 
-
\big( \gamma x\_n^* + \frac{\omega p_I}{(\omega+\gamma)}(\hat F_I x\_n^*) + \hat F_I 
\big).
$$
After simplification, we use the fact that the following expression is positive
$$(\frac{(\omega+\gamma)}{(\omega p_I+\gamma)}) (1+ \frac{\omega p_I x\_n^*}{(\omega+\gamma)}) -1. $$
This is because $(\frac{(\omega+\gamma)}{(\omega p_I+\gamma)})>1$ and  $(1+ \frac{\omega p_I x\_n^*}{(\omega+\gamma)}) >1$.

(b) $\pro (B/C)>0$;
% %%%%%%%
\subsection{rate of returning tests} 
\ali{needs editig}
The linearization of $\Rnum$ around $\rho=0$ is
\begin{equation}\label{linearization}
\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(-\theta\_w)(\gamma w_S+\omega w_I) + (-\theta\_c)p_I w_I \omega^2 \Big). 
\end{equation}
So when $\rho \approx 0$ we have $$\partial{\Rnum}/\partial{\omega} \approx  \frac{-\beta \rho}{\gamma w_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),$$
where $a=(-\theta\_w)w_I-(-\theta\_c)p_I w_I$, $b=2(-\theta\_w)\gamma w_S$ and $c=(-\theta\_w)\gamma^2 w_S$.

@@

\subsection{Expensive vs.\ cheap tests}

>>>>

It remains hard to show the reduction of $\Rnum$ with respect to \percap testing intensity, $\rho$, and the speed of the test return, $\omega$, for the feasible ranges of the parameters, i.e., $0<\omega$, $0 \leq \rho<\omega$, and $0 \leq \theta\_w \leq \theta\_c \leq 1$. One easy way to do such an analysis is based on the fact that $\rho$ is \percap rate so for a large population size, $N_0 \gg 1$, $\rho$ is very small or close to 0, comparing to $N_0$. This provides a base to linearly approximate $\Rnum$ when $\rho$ is close to 0, and use this approximation to analyze the behaviour of $\Rnum$ with respect to $\omega$ (see section \ref{taylor}). In the next section we provide an equivalent representation of $\Rnum$ for our model to provide to follow proving that more testing intensity decreases $\Rnum$ for a general rane of parameters.

% %%%%%%%
\subsection{More testing intensity decreases $\Rnum$}

This is to prove that $\pro \Delta > 0$ with $\Delta$ defined in Eq. \eqref{eq:abc}, and thus $\pro \Rnum < 0$, where $\Rnum$ is the basic reproduction number, given in Eqs. \eqref{R0}.

% %%%%%%%
\subsection{rate of returning tests} \label{taylor}
\ali{needs editig}
The linearization of $\Rnum$ around $\rho=0$ is
\begin{equation}\label{linearization}
\Rnum \approx \beta/\gamma + \frac{\beta \rho}{\omega (\omega+\gamma) \gamma^2 w_S} \Big(\gamma(-\theta\_w)(\gamma w_S+\omega w_I) + (-\theta\_c)p_I w_I \omega^2 \Big). 
\end{equation}
So when $\rho \approx 0$ we have $$\partial{\Rnum}/\partial{\omega} \approx  \frac{-\beta \rho}{\gamma w_S\omega^2 (\gamma+\omega)^2}  (a \omega^2 + b \omega + c),$$
where $a=(-\theta\_w)w_I-(-\theta\_c)p_I w_I$, $b=2(-\theta\_w)\gamma w_S$ and $c=(-\theta\_w)\gamma^2 w_S$.

@@

% %%%%%%%
\subsection{Expensive vs.\ cheap tests}
